\documentclass[letterpaper,12pt]{article}
\input{preamble}

% generics line 5: title, line 6: topic, line 7: date
\title{Fields}
\topic{Condensed Matter}
\date{Jul -- 14 -- 2024}



\author{Alberto Ruiz-Biestro}

\begin{document}
\maketitle
\tableofcontents
\thispagestyle{empty}

\clearpage\pagestyle{fancy}
\section{Calculus of variations}

\subsection{Functional Derivative}

\mrem{Definitions}{%
  \begin{outline}
    \1 \textbf{Field}: a map from a base manifold $\mcM$ into a target space $T$.
    \1 \textbf{Functional}: assigns each $\varphi$ a number $S[\varphi]$, i.e. it maps a field onto $\bbR$. Formally, it is defined through the map
    \[J:C^\infty(\bbR)\rightarrow\bbR\]
    \[\qq{where}J[y]=\int_{x_1}^{x_2}\dd{x}\,f(x,y,y',y'',\cdots,y^{(n)})\]
  \end{outline}
}

\begin{table}[h!]
  \centering
  \begin{tabular}{l l l}
    \hline\hline
    Entity & Discrete & Continuous\\ \hline
    Argument & vector $\vb{f}$ & function $f$\\
    Differential & $\dd{F_{\vb{f}}}(\vb{g})$ & $\delta F_{f}(g)$\\
    Cartesian basis & $\vb{e}_n$ & $\delta_x$\\
    Scalar product & \(\sum_nf_ng_n\) & \(\int\dd{x}f(x)g(x)\) \\
    Derivative & \(\pdv*{f_n}F(\vb{f})\) & \(\delta F[f]/\delta f(x)\) \\\hline\hline
  \end{tabular}
\end{table}

A functional derivative (also called a \emph{Frechet} derivative) is the generalization of a directional derivative $\pdv*{J}{y_i}$ for a continuum of coordinates $y_i\rightarrow y(x)$. Similar to multivariate calculus, finding the function that gives the extrema of $J[y]$ amounts to setting its functional derivative to zero.

\begin{equation}
\delta J=\sum_{i}\pdv{J}{y_i}\delta y_i\rightarrow\int_{x_1}^{x_2}\dd{x}\qty(\frac{\delta J}{\delta y(x)})\,\delta y(x)
\end{equation}

For the general case where $f$ depends on $n$ derivatives of $y(x)$, one has
\begin{equation}
  \frac{\delta J}{\delta y(x)}=\pdv{f}{y}-\dv{x}\qty(\pdv{f}{y'})
  +\dv[2]{x}\qty(\pdv{f}{y''})-\dv[3]{x}\qty(\pdv{f}{y'''})+\cdots=0.
\end{equation}
Relating the functional derivative to partial derivatives is essential for analyzing problems. In Lagrangian mechanics, the action functional and the functional derivative take the form of
\mnote{Lagrangian functional}{%
\begin{equation}
  S[q]=\int_{t_0}^{t_f}\dd{t}L(t,q^i,{q'}^{i}),\qquad \frac{\delta L}{\delta q(t)}=\dv{t}\qty(\pdv{L}{\dot{q}})-\pdv{L}{q}=0
\end{equation}
}


\subsection{Continuous systems}

Continuum systems and field theories require infinite degrees of freedom. For notational purposes, we define
\begin{equation}
  \boxed{\varphi_\mu\equiv\pdv{\varphi}{x^\mu}}.
\end{equation}
The Lagrangians evaluated over a manifold $\mcM$ will usually take the form of
\begin{equation}
  \boxed{S[\varphi]=\int_{\mcM}\dd{t}L=\int_{\mcM}\dd[m]{x}\mcL(x^\mu,\varphi,\varphi_\mu)
  }
\end{equation}

\mnote{Vanishing at boundary}{We assume the field variation vanishes at the boundary, $\theta\big|_{\partial\mcM}=\delta\varphi\big|_{\partial\mcM}=0$. This eliminates any boundary integrals when integrating by parts. However, this vanishing need not be always the case.}


\mrem{Divergence theorem}{Using the established notation, the divergence theorem looks like this
  \[\int_{\Omega}\dd[m]{x}\qty(\pdv{A^\mu}{x^\mu})=\int_{\partial \Omega}\dd{S}A^\mu n_\mu\]
}

\mnote{First Principles}{In continuum problems, it is often easier to 
derive the equations of motion from \emph{first principles}, rather 
than using the continuum Euler-Lagrange equations,
\begin{equation}
  \frac{\delta S}{\delta \varphi^i(x)}=\pdv{\mcL}{\varphi^i(x)}-\pdv{x^\mu}\qty(\pdv{\mcL}{\varphi^i_\mu(x)}).
\end{equation}
Variating by hand will usually require the following identity, 
\[\boxed{\delta(\varphi_\mu(x))=\pdv{x^\mu}\delta\varphi(x)}.\]
}

\subsection{Noether's Theorem}
\mrem{Continuum Noether's Theorem}{A continuous symmetry entails a classically conserved current.}

\mnote{Momenta}{%
  \begin{outline}
    Imagine a wave on a string with fixed ends. The distinction between momentum and pseudo-momentum is appreciated by analyzing the associated symmetry,
    \1 \(\qq{Pseudo momentum}\Leftrightarrow \qq{invariance under}y(x)\rightarrow y(x-a)\qq{(loc. of wave)}\)
    \1 \(\qq{Momentum}\Leftrightarrow \qq{invariance under translation of whole system}\)
  \end{outline}
}

\mtip{Curl}{Curl curl identity 
\[\curl(\frac{1}{2})\]}

\subsection{Electrodynamics}
More fundamental than Maxwell's equations, for modern electrodynamics, is the concept of \textbf{symmetry}; an electromagnetic theory must be Lorenz invariant. We restrict our analysis to vacuum theory, where $\vb{E}=\vb{D}$ and $\vb{B}=\vb{H}$.

The Lagrangian (or \emph{action density}) that gives rise to Maxwell's equations is
\begin{equation}
  \boxed{\mcS=(\varphi,\varphi_\mu)=\frac{1}{2}\qty(\dot{\vb{A}}\cdot{\dot{\vb{A}}}-\abs{\curl{A}}^2)}.
\end{equation}
The terms follow a similar relation to $T-U$. Should there be an external current, the Lagrangian includes an additional source term

% TODO: axial gauge

% the axial gauge does not completely fix the gauge
\mnote{Gauge fixing}{It means removing the freedom one has in specifying the four-vector potential $A_\mu$, i.e. the axial-gauge and Lorenz gauge, $n_\mu{A^\mu}=0$ and $\partial_\mu A^\mu$, do not completely fix the gauge.}

\mrem{Contravariant and Covariant}{
Consider a change of basis from $\vb{e}\rightarrow\vb{e}'$, where $\vb{e},\vb{e}'\in V$, given by
\[\boxed{\vb{e}_\nu=a_\nu^\mu\vb{e}'_\mu}.\]
\centering
    \begin{tabular}{l | c | c}
      \hline\hline
      & \textbf{Covariant} & \textbf{Contravariant}\\
      \hline
      Elements & covectors $\vb{f}\in V^*$ & vectors $\vb{x}\in V$\\
      Transformation & $f_\nu=a_\nu^\mu {f'}_\mu$ & ${x'}^\mu=a_\nu^\mu x^\nu\qq{i.e.} x^\nu=(a_\nu^\mu)^{-1}{x'}^\mu$\\
      Derivative & $\partial_{\mu}=(+\partial_t,\grad)$ & $\partial^{\mu}=(-\partial_t,\grad)$
    \end{tabular}


}

\section{Quantizing fields}

\end{document}
